---
title: "USGS Code Sample"
author: "Daphne Virlar"
date: "6/10/2020"
output: html_document
---

```{r, include=FALSE}
# The following set of html code shows I am familiar with html.I frequently use html to change the layout of my markdown files to make my assignments and work products more aesthetically pleasing to view.
```

<style type="text/css">

h1.title {
  text-align: center;
  font-family: "Times New Roman", Times, serif;
}
h4.author {
  font-family: "Times New Roman", Times, serif;
  text-align: center;
  font-size: 26px;
}
h4.date { 
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}

body {
font-family: "Times New Roman", Times, serif;
font-size: 20px;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load necessary libraries
library(tidyverse)
library(extrafont) # allows for font changes within ggplot

library(vcdExtra) # this allows for the expand.dft function in example 2, where I change frequency values into individual data points

library(ncdf4) # allows for the import of climate data, will be used in example 3
library(patchwork) # allows user to put multiple plots together, will be used in example 3
library(chron) # allows user to manage time and date data, will be used in example 3
library(lubridate) # allows user to manage time and date data, will be used in example 3
```

<br>

Example 1: Carapice length of lobsters in the Santa Barbara Channel within different LTER sites

```{r, echo=FALSE, warning=FALSE, fig.align='center'}
# Data was retrieved from: https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-sbc.77.4


# Read in the data
lobster_abundance <- suppressMessages(read_csv("Lobster_Abundance_All_Years.csv")) # data frame with lobster abundance data


# Tidy up the data
lobster_abundance_tidy <- lobster_abundance %>% 
  as.data.frame(lobster_abundance_tidy) %>%
  filter(COUNT != "0") %>%
  expand.dft(freq = "COUNT") %>% 
  filter(SIZE_MM >= "0") # It's impossible to have a negative body size, so I am filtering out those data points


# Create new data frame for MPA only data, and data only from 2012 and 2019
MPA12_19 <- lobster_abundance_tidy %>% 
  filter(YEAR == "2012" | YEAR == "2019") %>% 
  mutate(PROTECTED = case_when(
    SITE == "IVEE" ~ "MPA",
    SITE == "NAPL" ~ "MPA",
    SITE == "AQUE" ~ "Non-MPA",
    SITE == "CARP" ~ "Non-MPA",
    SITE == "MOHK" ~ "Non-MPA")) %>% 
  mutate(SITE = case_when(
    SITE == "IVEE" ~ "Isla Vista",
    SITE == "NAPL" ~ "Naples",
    SITE == "AQUE" ~ "Arroyo Quemado",
    SITE == "CARP" ~ "Carpinteria",
    SITE == "MOHK" ~ "Mohawk"
  ))

# Graph the size distributions of lobsters at LTER sites in the Santa Barbara Channel
MPA_Graph <- ggplot(MPA12_19, aes(x= SITE, y= SIZE_MM))+
  geom_boxplot(aes(fill = PROTECTED))+
  facet_wrap(~ YEAR)+
  theme_classic()+
  scale_x_discrete(limits = c("Arroyo Quemado", "Carpinteria", "Mohawk", "Isla Vista", "Naples"))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 90, vjust=0.6))+
  theme(text=element_text(family="Times New Roman"))+
  ylab("Carapace Length (mm)")+
  xlab("Site")+
  ggtitle("Size Distribution of Lobsters at Santa Barbara Channel LTER Sites")+
  scale_fill_discrete(name = "Protected Status")


MPA_Graph + theme(panel.spacing = unit(3, "lines")) 

```

<br>
<br>

Example 2: Graphing predicted snowfall in Colorado, using open-source climate data from UCAR.

```{r, echo=FALSE, fig.align='center'}
# Data was retrieved from: http://www.cesm.ucar.edu/projects/community-projects/LENS/data-sets.html


# Read in data
en_01 <- nc_open("b.e11.BRCP85C5CNBDRD.f09_g16.001.clm2.h0.SNOW.200601-208012.nc")
en_02 <- nc_open("b.e11.BRCP85C5CNBDRD.f09_g16.002.clm2.h0.SNOW.200601-208012.nc")
en_03 <- nc_open("b.e11.BRCP85C5CNBDRD.f09_g16.003.clm2.h0.SNOW.200601-208012.nc")


# Set the latitude and longitude for Colorado
lat <- ncvar_get(en_01, "lat")
lon <- ncvar_get(en_01, "lon")

# Zoom in on Colorado
lats = which(lat >= 37 & lat <= 41)
lons = which(lon >= 251 & lon <= 256)
  # the longitudes have already been subtracted from 360, per the requirements of the model
  # Used Denver as the Eastern cut off for the state


# Set the time frame
begin_time="1/15/2006"
end_time="12/15/2080"


# Combine dates into a sequence
all_time <- seq.dates(begin_time, end_time, by="months")


# Convert snowfall from mm/s to in/month
snow_01 <- ncvar_get(en_01, "SNOW") * (1/25.4) * (2.628e+6)
snow_02 <- ncvar_get(en_02, "SNOW") * (1/25.4) * (2.628e+6)
snow_03 <- ncvar_get(en_03, "SNOW") * (1/25.4) * (2.628e+6)


# Calculate Regional Average
snow1_regavg <- apply(snow_01[lons,lats,],3,sum)
snow2_regavg <- apply(snow_02[lons,lats,],3,sum)
snow3_regavg <- apply(snow_03[lons,lats,],3,sum)


# Combine Time and Regional Average
snow1_data <- data.frame(all_time, snow = round(snow1_regavg, 2))
snow2_data <- data.frame(all_time, snow = round(snow2_regavg, 2))
snow3_data <- data.frame(all_time, snow = round(snow3_regavg, 2))


# Add a month column for all climate ensembles
snow1_data$month = month(all_time)
snow2_data$month = month(all_time)
snow3_data$month = month(all_time)


# Create data frame that identifies months within a ski season, taking into account that January of one year is part of the same ski season for December of the previous year. Repeat this for all three climate ensembles.
snow_season_01 <- mutate(snow1_data, 
                   season_year = ifelse(month(all_time) == 1, year(all_time)-1, 
                                 ifelse(month(all_time) == 2, year(all_time)-1, 
                                 ifelse(month(all_time) == 3, year(all_time)-1,
                                 ifelse(month(all_time) == 4, year(all_time)-1,
                                 ifelse(month(all_time) == 5, year(all_time)-1,
                                        year(all_time)))))),
       season = case_when(
        month(all_time) %in% c(11, 12, 1, 2, 3, 4, 5) ~ "Ski_On",
        month(all_time) %in% c(6, 7, 8, 9, 10) ~ "Ski_Off",
        T ~ NA_character_
       ))


snow_season_02 <- mutate(snow2_data, 
                   season_year = ifelse(month(all_time) == 1, year(all_time)-1, 
                                 ifelse(month(all_time) == 2, year(all_time)-1, 
                                 ifelse(month(all_time) == 3, year(all_time)-1,
                                 ifelse(month(all_time) == 4, year(all_time)-1,
                                 ifelse(month(all_time) == 5, year(all_time)-1,
                                        year(all_time)))))),
       season = case_when(
        month(all_time) %in% c(11, 12, 1, 2, 3, 4, 5) ~ "Ski_On",
        month(all_time) %in% c(6, 7, 8, 9, 10) ~ "Ski_Off",
        T ~ NA_character_
       ))


snow_season_03 <- mutate(snow3_data, 
                   season_year = ifelse(month(all_time) == 1, year(all_time)-1, 
                                 ifelse(month(all_time) == 2, year(all_time)-1, 
                                 ifelse(month(all_time) == 3, year(all_time)-1,
                                 ifelse(month(all_time) == 4, year(all_time)-1,
                                 ifelse(month(all_time) == 5, year(all_time)-1,
                                        year(all_time)))))),
       season = case_when(
        month(all_time) %in% c(11, 12, 1, 2, 3, 4, 5) ~ "Ski_On",
        month(all_time) %in% c(6, 7, 8, 9, 10) ~ "Ski_Off",
        T ~ NA_character_
       ))




# Sum the total ski season snowfall from 2020 to 2080 for each climate ensemble.
snow_season_01_summary <- snow_season_01 %>%
  filter(season == "Ski_On") %>% 
  filter(season_year > 2019) %>% 
  filter(season_year< 2080) %>% 
  group_by(season_year) %>% 
  summarize(total_snow = sum(snow))


snow_season_02_summary <- snow_season_02 %>%
  filter(season == "Ski_On") %>%
  filter(season_year > 2019) %>% 
  filter(season_year< 2080) %>% 
  group_by(season_year) %>% 
  summarize(total_snow = sum(snow))


snow_season_03_summary <- snow_season_03 %>%
  filter(season == "Ski_On") %>% 
  filter(season_year > 2019) %>% 
  filter(season_year< 2080) %>% 
  group_by(season_year) %>% 
  summarize(total_snow = sum(snow))


# Graph the time series of total predicted snowfall per ski season from 2020-2080, for each climate ensemble. Also, add in the predicted average high and low snowfall of 212.51 in/month and 133.21 in/month respectively

en_01_graph <- ggplot(snow_season_01_summary, aes(x = season_year, y = total_snow)) +
  geom_line(color="goldenrod") +
  geom_hline(yintercept = 212, color="#C70000", linetype="dashed")  +
  geom_hline(yintercept = 133, color="#009BB0", linetype="dashed") +
  scale_y_continuous(limits=c(0,300), expand=c(0,0)) +
  xlab("Year \n") + 
  ylab("Total Snow \n(in)") +
  ggtitle("Predicted Snowfall per Season (2020-2080) Under RCP 8.5 Ensemble 1") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Times New Roman"))+
  theme(plot.margin = unit(c(1,1,2,1), "points"))


en_02_graph <- ggplot(snow_season_02_summary, aes(x = season_year, y = total_snow)) +
  geom_line(color="#7A5028") + 
  geom_hline(yintercept = 212, color="#C70000", linetype="dashed")  +
  geom_hline(yintercept = 133, color="#009BB0", linetype="dashed") +
  scale_y_continuous(limits=c(0,300), expand=c(0,0)) +
  xlab("Year \n") + 
  ylab("Total Snow \n(in)") +
  ggtitle("\nPredicted Snowfall per Season (2020-2080) Under RCP 8.5 Ensemble 2") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Times New Roman"))+
  theme(plot.margin = unit(c(1,1,2,1), "points"))


en_03_graph <- ggplot(snow_season_03_summary, aes(x = season_year, y = total_snow)) +
  geom_line(color="#114C54") +
  geom_hline(yintercept = 212, color="#C70000", linetype="dashed")  +
  geom_hline(yintercept = 133, color="#009BB0", linetype="dashed") +
  scale_y_continuous(limits=c(0,300), expand=c(0,0)) +
  xlab("Year \n") + 
  ylab("Total Snow \n(in)") +
  ggtitle("\nPredicted Snowfall per Season (2020-2080) Under RCP 8.5 Ensemble 3") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Times New Roman"))


wrap_plots(en_01_graph / en_02_graph / en_03_graph)

```